### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
Names rmse figures

---
title: "Poster figures"
output: 
  html_document:
    self_contained: false  
date: "2024-11-11"
---

```{r setup, include=FALSE, cache = FALSE}
library(lubridate)
library(ggplot2)
library(data.table)
library(metR)
library(rcdo)
library(tagger)
library(patchwork)

knitr::opts_chunk$set(cache = TRUE)

cdo_options_set("-L")
```

```{r helpers, cache = FALSE}
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))
```

```{r}

nb_yellow <- gglgbtq::palette_lgbtq("nonbinary")[1] |> 
  colorspace::darken(0.3)
nb_purple <- gglgbtq::palette_lgbtq("nonbinary")[3]
colours_models <- c(
  S2 = nb_purple,
  S1 = nb_yellow,
  cdr = "black",
  persistence = "black",
  climatology = "gray50"
)

labels_models <- c(
  S2 = "ACCESS-S2",
  S1 = "ACCESS-S1",
  cdr = "CDR",
  bt = "Bootstrap",
  era5 = "ERA5",
  osi = "OSI",
  persistence = "Persistence",
  climatology = "Climatology"
)

scale_color_models <-   ggplot2::scale_color_manual(NULL,
                                                    values = colours_models,
                                                    labels = labels_models
)

scale_fill_models <-   ggplot2::scale_fill_manual(NULL,
                                                  values = colours_models,
                                                  labels = labels_models
)

trans_pink <- gglgbtq::palette_lgbtq("transgender")[2] |> 
  colorspace::darken(0.2)
trans_blue <- gglgbtq::palette_lgbtq("transgender")[1] |> 
  colorspace::darken(0.2)

```

```{r datasets}
datasets <- list(cdr = CDR())
which_dataset <- "cdr"
```

### Bias

```{r extent_daily}
extent_daily <- datasets |>
  lapply(extent) |>
  lapply(\(x) ReadNetCDF(x, "aice")) |>
  rbindlist(idcol = "dataset") |>
  _[, let(lon = NULL, lat = NULL)]
```

```{r extent_climatology}
extent_climatology <- extent_daily |>  
  copy() |>
  _[aice == 0, aice := NA] |> 
  _[!(month(time) == 2 & mday(time) == 29)] |> 
  _[, time := update(time, year = 2001)] |> 
  _[, median_ci(aice), by = .(time, dataset)]

```

```{r hindcast_extent}
hindcast_extent <- fread(here::here("data/derived/hindcast_extent.csv")) |> 
  _[, model := relevel(factor(model), "S2")] |> 
  _[, lag := as.numeric(as.Date(time) - as.Date(forecast_time))] |> 
  _[lag > 0] |> 
  _[!(month(forecast_time) == 1 & member == 7 & model == "S1")]  # This member is wrong
```

```{r}
cm <- 2.54
```

```{r fig-hindcast-extent, fig.height=11.99/cm, fig.width=23.59/cm}
#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  # _[]
  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset], 
            aes(x = as.Date(time),
                group = dataset)) + 
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
             aes(fill = month), 
             shape = 24, size = 2.2) +

  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
             shape = 21, size = 2) +

  scale_x_date(NULL, date_breaks = "1 month", 
               date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = labels_extent) +
  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
                     aesthetics = c("fill", "color"), guide = "none") +
  facet_wrap(~ model, ncol = 1) +
  tag_facets() 
```

```{r hindcast_extent-bias, fig.height=11.99/cm, fig.width=23.59/cm}
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  _[copy(extent_climatology)[, time := as.Date(time)], on = c(time2 = "time")] |> 
  _[] |> 
  ggplot(aes(time2, mid - i.mid)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
             aes(fill = month), 
             shape = 24, size = 2.2) +

  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
             shape = 21, size = 2) +

  scale_x_date(NULL, date_breaks = "1 month", 
               date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = labels_extent) +
  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
                     aesthetics = c("fill", "color"), guide = "none") +
  facet_wrap(~ model, ncol = 1) +
  tag_facets() 
```

```{r extent-delta, fig.height=11.99/cm, fig.width=24.09/cm}

dif_hindcast <- hindcast_extent |> 
  copy() |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]

extent_daily |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
  _[dataset == which_dataset] |> 
  ggplot(aes(time, mid)) +
  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
  geom_smooth(aes(group = dataset, color = dataset),
              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
  geom_smooth(data = dif_hindcast,
              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
  scale_x_datetime(NULL, date_breaks = "1 month", 
                   date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = \(x) labels_extent(x, units = "M km²/day")) +
  theme(legend.position = "inside", legend.position.inside = c(0.9, 0.9), 
        legend.direction = "vertical") +
  scale_color_models
```

```{r hindcast_clim}
zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
lt$year*12 + lt$mon } 
month_differnece <- function(d1, d2) { monnb(d2) - monnb(d1) }

month_diff <- function(month1, month2) {
  d <- (month2 - month1)

  (d %% 12)

}

months <- 1:12

hindcast_clim <- purrr::pmap(CJ(month = 1:12, version = c("S2", "S1")), \(month, version) {
  here::here(glue::glue("data/derived/climatology/{version}/{zeropad(month)}/em.nc")) |> 
    # cdo_selmonth(month) |> 
    cdo_monmean() |> 
    cdo_execute() |> 
    ReadNetCDF("aice") |> 
    _[, let(forecast_month = month, 
            version = version)] 
}) |> 
  rbindlist() |> 
  _[, forecast_month := as.numeric(forecast_month)] |> 
  _[, forecast_time := update(time, month = forecast_month)] |> 
  _[, time := as.Date(update(time, day = 1))]

obs_clim <- CDR() |> 
  climatology() |> 
  cdo_ymonmean() |>
  cdo_execute() |> 
  ReadNetCDF(c(obs = "aice")) |> 
  _[, time := as.Date(update(time, day = 1))]

hindcast_clim <- hindcast_clim |> 
  merge(obs_clim) |> 
  _[, month := month(time)] |>
  _[, month_diff := month_differnece(time[1], forecast_time[1]), by = .(forecast_month, time)]

hindcast_clim[, month_diff := month_diff(forecast_month[1], month[1]), by = .(forecast_month, month)]

```

```{r bias_plot}
#| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.

bias_plot <- function(hindcast_clim, which_version, which_lead) {

  breaks <- seq(-1, 1, by = 0.1)
  breaks <- breaks[breaks != 0]
  hindcast_clim |> 
    _[version == which_version] |>
    _[month_diff == which_lead] |> 
    # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
    # setnames(which_dataset, "obs") |>
    ggplot(aes(xgrid, ygrid)) +
    geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
                      breaks = breaks) +
    # geom_contour2(aes(z = aice - obs), breaks = AnchorBreaks(binwidth = 0.1, exclude = 0),
    #               linewidth = 0.1) +
    scale_fill_divergent_discretised(paste0(which_version, " bias at ", which_lead, " months at lead time"),
                                     low = trans_pink,
                                     high = trans_blue) +
    geomcoord_antarctica +
    geom_antarctica_fill +
    facet_wrap(~ month, 
               labeller = labeller(month = setNames(1:12, month.abb)),
               nrow = 2) +
    wide_legend +
    tag_facets()
}

```

```{r}
width_bias <- 22.15/cm
height_bias <- 12.3/cm
```

```{r bias_S2_0, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}

hindcast_clim |> 
  bias_plot("S2", 0)
```

```{r bias_S1_0, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}

hindcast_clim |> 
  bias_plot("S1", 0)
```

lead
```{r bias_S2_6, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}
hindcast_clim |> 
  bias_plot("S2", 6)

```

```{r bias_S1_6, fig.width=width_bias, fig.height=height_bias, dependson="bias_plot"}
hindcast_clim |> 
  bias_plot("S1", 6)

```

### Anomalies

```{r fig-extent-anom, fig.height=11.99/cm, fig.width=23.22/cm}
#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).

months_difference <- function(x, y) {
  lubridate::interval(y, x) %/% months(1)
}

monthly_extent <- extent_daily |> 
  _[dataset == which_dataset] |> 
  _[aice == 0, aice := NA]  |> 
  _[, .(aice = mean(aice)), by = .(dataset, time = as.Date(round_date(time, "month")))] |>
  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE), 
    by = .(month(time), dataset)]

N <- uniqueN(monthly_extent$time)

s1_range <- hindcast_extent[model == "S1"][, range(time)]
s2_range <- hindcast_extent[model == "S2"][, range(time)]

hindcast_extent |> 
  copy() |> 
  _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]), 
    by = .(lag, model, time = update(time, year = 2000))] |> 
  _[, .(forecast = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
  _[, lag := months_difference(time, forecast_time)] |> 
  _[, merge(.SD, monthly_extent, all = TRUE), by = .(model, lag)] |> 
  _[lag %in% (c(1, 3, 5, 7) - 1)] |>
  _[time %between% s2_range] |> 
  ggplot(aes(time)) +
  geom_line(aes(y = aice, group = dataset, color = dataset)) +
  geom_line(aes(y = forecast, color = model))  +
  geom_text(data = \(x) x[time %between% s1_range, 
                          .(text = sqrt(mean((aice - forecast)^2, na.rm = TRUE))), by = .(lag, model)],
            aes(label = paste0(scales::number(text, scale = 1e-12), " M km²")), x = -Inf, y = Inf,
            vjust = 1.2, hjust = -0.2) +
  scale_y_continuous(NULL,
                     labels = labels_extent) +
  scale_x_date(NULL, expand = c(0, 0)) +
  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Month: ", x))) +
  scale_color_models

```

### RMSE

```{r errors}
errors <- rbind(
  readRDS(here::here("data/derived/rmse.Rds")),
  readRDS(here::here("data/derived/iiee.Rds"))
) |> 
  _[value < 0.01, value := NA] |> 
  _[value > 3e13, value := NA] |> 
  _[, time := as.Date(time)] |> 
  _[, lag := as.numeric(time - time_forecast)] |> 
  _[lag > 0] |> 
  _[!(month(time_forecast) == 1 & member == "07" & version == "S1")]  # This member is wrong

```

```{r clim_std}
clim_std <- datasets |> 
  lapply(\(x) x |> 
           anomalies() |> 
           cdo_sqr() |> 
           cdo_fldmean() |> 
           cdo_sqrt() |> 
           cdo_execute() |> 
           ReadNetCDF(c(climatology = "aice")) |> 
           _[, let(lat = NULL, lon = NULL)]
  ) |> 
  rbindlist(idcol = "obs_dataset") |> 
  _[, time := as.Date(time)] |> 
  _[climatology == 0, climatology := NA] |> 
  _[climatology < 0.05, climatology := NA]

```

```{r}
clim_std2 <- clim_std[errors, on = .NATURAL] |> 
  _[, let(climatology = NULL,
          value = climatology)] |> 
  na.omit() |> 
  _[, version := "climatology"]
```

```{r fig-rmse}
#| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies. 
#| fig.height: 5
#| fig.width: 9

month_from_lag <- function(day = 1, step = 2) {
  force(day)
  force(step)
  function(x) {
    unique(x[, .(lag, month)])[, time := make_date(month = month, day = 1) + lag] |> 
      _[mday(time) == day, .(lag, month, 
                             month2 = lubridate::month(time, label = TRUE)
      )] |> 
      _[, .SD[((seq_len(.N) + 1) %% step) == 0], by = month]
  }
}

g <- clim_std2 |> 
  rbind(errors) |>
  _[obs_dataset == which_dataset] |> 
  _[measure == "rmse"] |>
  _[, time2 := update(time, year = 2000)] |> 
  _[, median_ci(value),
    by = .(lag, version, measure, month(time_forecast), obs_dataset)] |>
  ggplot(aes(lag, mid)) +

  geom_vline(data = month_from_lag(1, 1), aes(xintercept = lag), 
             colour = "gray50", linewidth = 0.1) +

  geom_text(data = month_from_lag(15, 1), aes(y = Inf, label = month2),
            size = 2.5,
            vjust = 1) +

  geom_text(data = month_from_lag(1, 1), aes(y = -Inf, label = lag), 
            size = 2.5,
            vjust = 0) +

  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version,
                  group = interaction(obs_dataset, version)),
              alpha = 0.1) +
  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +

  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(y = "RMSE", 
       x = "Lead time") +
  scale_color_models +
  scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
```

```{r}
width_errors <- 19.43/cm
height_errors <- 11.99/cm
```

```{r rmse_persistence, fig.width=width_errors, fig.height=height_errors}
g %+% g$data[version != "climatology"]
```

```{r rmse_climatology, fig.width=width_errors, fig.height=height_errors}
g %+% g$data[version != "persistence"]
```

```{r}
# ACHTUNG: we have something going on with the 29th of february.

clim_std2 |> 
  rbind(errors) |> 
  _[measure == "rmse"] |> 
  _[obs_dataset == which_dataset] |> 
  _[, .(rmse = median(value, na.rm = TRUE)), 
    by = .(time = update(time, year = 2001), 
           time_forecast = update(time_forecast, year = 2001), 
           lag,
           version)] |> 
  # _[version == "climatology"] |>
  ggplot(aes(time, rmse)) +
  geom_line(data = \(x) x[version == "climatology"],
            aes(group = interaction(time_forecast, month(time_forecast) <= month(time)))) +
  # geom_line(data = \(x) x[version == "S2"],
  #           aes(color = factor(month(time_forecast)),
  #               group = interaction(time_forecast, month(time_forecast) <= month(time)))) +
  # geom_point(data = \(x) x[version == "S2" & lag ==1], aes(color =  factor(month(time_forecast)))) +
  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
                     aesthetics = c("fill", "color"), guide = "none") +
  scale_y_continuous(limits = c(0, NA)) +
  geom_vline(xintercept = as.Date("2001-02-28"))
```

```{r error_zoom, fig.width=width_errors, fig.height=height_errors}
p <- g +
  coord_cartesian(xlim = c(0, 100))

p$layers <- p$layers[-4]
p
```

```{r}
topo_lonlat <- rcdo::cdo_topo(grid = "r180x90") |>
  rcdo::cdo_execute(options = c("-f nc")) |>
  metR::ReadNetCDF(c(z = "topo"), subset = list(lat = c(-90, -60))) |> 
  ggperiodic::periodic(lon = c(0, 360)) 
```

```{r rmse_lon_mean}
rmse_lon_mean <- here::here("data/derived/rmse_lon.Rds") |> 
  readRDS() |> 
  _[version != "S1"] |> 
  _[, lag := as.numeric(as.Date(time) - time_forecast)] |> 
  _[, base := value[version == "persistence"], 
    by = .(time, lon, lag, obs_dataset)] |> 
  _[version != "persistence"] |>
  _[, .(dif = median(value - base, na.rm = TRUE)),
    by = .(lag, lon, version, obs_dataset, month = factor(month(time_forecast)))]
```

```{r fig-rmse_lon, fig.height=11.99/cm, fig.width=23.1/cm}
#| fig.cap: Meadian difference between RMSE of +S2 forecasts and persistence forecast at longitudinal bands.

topo_inset <- list(topo_lonlat) |>
  rep(4) |>
  rbindlist(idcol = "month") |>
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = z),
                    breaks = c(0, Inf), fill = "gray90") +
  geom_contour2(aes(z = z),
                breaks = 0) +
  facet_wrap(~month, labeller = labeller(mohth = rep("", 4)), nrow = 1) +
  scale_y_latitude(limits = c(-90, -60)) +
  scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +
  theme_void() +
  theme(strip.background = element_blank(),
        strip.text = element_blank(),
        aspect.ratio = 0.1,
        plot.margin = margin(c(-50, 0, 0, 0)))

theme_margin <- theme_get()$plot.margin
theme_margin[3] <- margin(-10)[1]

binwidth <- 0.02
rmse_lon_mean |> 
  _[obs_dataset == which_dataset] |> 
  # _[month %in% 1:4] |> 
  ggperiodic::periodic(lon = c(0, 360)) |> 
  ggplot(aes(lon, lag)) +
  geom_contour_fill(aes(z = dif, fill = after_stat(level)),
                    breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth))  +
  geom_contour_tanaka(aes(z = dif), range = c(0.01, 0.4),
                      breaks = AnchorBreaks(0, exclude = 0, binwidth = binwidth)) +

  geom_hline(data = month_from_lag(1, 1), aes(yintercept = lag),
             colour = "gray50", linewidth = 0.1) +

  geom_text(data = month_from_lag(15, 1), aes(x = 361, label = month2),
            size = 2.5, 
            hjust = 0) +

  geom_text(data = month_from_lag(1, 1), aes(x = -2, label = lag),
            size = 2.5,
            hjust = 1) +
  # scale_fill_divergent_discretised() +
  scale_fill_divergent_discretised(high = trans_pink,
                                   low  = trans_blue) +
  scale_x_longitude(ticks = 60, expand = c(0.08, 0)) +

  scale_y_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(fill = NULL, 
       y = "Lead time") +
  wide_legend +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) +
  theme(panel.background = element_blank(), 
        tagger.panel.tag.background = element_rect(colour = "white", fill = "white"), 
        plot.margin = theme_margin) +
  tag_facets(position = list(x = 0.08, y = 1, hjust = 0, vjust = 1)) +

  topo_inset +

  plot_layout(ncol = 1, guides = "collect", heights = c(10, 1), axes = "collect")
```