### 2025-10-09 analysis/paper/access-ice.tex
Renders

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-mathphys-num,
  referee]{sn-jnl}

\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\usepackage[section]{placeins}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Evaluating the Importance of Initial Conditions for Antarctic Sea Ice Seasonal Predictability with a Fully Coupled Model},
  pdfauthor={Elio Campitelli; Ariaan Purich; Julie Arblaster; Eun-Pa Lim; Matthew Wheeler; Phillip Reid},
  pdfkeywords={sea ice, predictability, initial
conditions, forecasting},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Evaluating the Importance of Initial Conditions for Antarctic Sea
Ice Seasonal Predictability with a Fully Coupled Model]{Evaluating the
Importance of Initial Conditions for Antarctic Sea Ice Seasonal
Predictability with a Fully Coupled Model}

% author setup
\author*[1,2]{\fnm{Elio} \sur{Campitelli}}\email{elio.campitelli@monash.edu}\author[1,2]{\fnm{Ariaan} \sur{Purich}}\author[1,2]{\fnm{Julie} \sur{Arblaster}}\author[3]{\fnm{Eun-Pa} \sur{Lim}}\author[3]{\fnm{Matthew} \sur{Wheeler}}\author[3]{\fnm{Phillip} \sur{Reid}}
% affil setup
\affil[1]{\orgdiv{School of Earth, Atmosphere and
Environment}, \orgname{Monash University, Australia,}}
\affil[2]{\orgdiv{Securing Antarctica's Environmental
Future}, \orgname{Monash University, Australia}}
\affil[3]{\orgdiv{Research, Bureau of Meteorology}}

% abstract

\abstract{Accurate Antarctic sea ice forecasts are crucial for climate
monitoring and operational planning, yet they remain challenging due to
model biases and complex ice-ocean-atmosphere interactions. The two
versions of the Australian Bureau of Meteorology's ACCESS seasonal
forecast system, ACCESS-S1 and ACCESS-S2, use identical model
configuration and differ only in their initial conditions; primarily in
that ACCESS-S2 does not assimilate sea-ice observations. This makes for
a good ``natural experiment'' to analyse the role of initial conditions
on Antarctic sea ice forecasts using a long run and a moderately large
ensemble with a fully coupled model. Our analysis reveals that both
systems experience an extended melt season and delayed growth phase.
This leads to a significant negative sea-ice extent bias which is
corrected only in ACCESS-S1 by the data assimilation system. The impact
of the initial conditions on forecast errors varies dramatically by
season: summer and autumn initial conditions (January-April) provide
predictive skill for up to three months, with February initial
conditions being particularly crucial. In contrast, winter forecasts of
the two systems are statistically indistinguishable after only two
weeks. Regional analysis of forecast skill suggests that this winter
predictability barrier is most dramatic over Eastern Antarctica, where
even ACCESS-S1 shows negative skill. These findings highlight the
critical importance of comprehensive year-round sampling in
predictability studies and suggest that operational sea-ice data
assimilation efforts should prioritize the summer-autumn period when
initial conditions have maximum impact on forecast skill.}

% keywords
\keywords{sea ice,  predictability,  initial conditions,  forecasting}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

Accurately modelling Antarctic sea ice is essential for understanding
processes and improving climate projections to inform adaptation
strategies. Accurate seasonal to sub-seasonal forecasts are also crucial
for operation contingency planning in and around the Antarctic
continent, including scientific missions, fisheries, and tourism
\citep{desilva2020, wagner2020}. Improvements in modelled sea-ice might
also help improve weather forecasts over and away from sea-ice regions
\citep{rinke2006, wang2024, semmler2016}.

However, Antarctic sea ice forecasts have lagged behind Arctic sea ice
forecasts due to model biases, inherent large variability and complexity
\citep{zampieri2019, gao2024a}. Dynamical seasonal forecasts of summer
Antarctic sea ice have been shown to perform worse than relatively
simpler statistical methods \citep{massonnet2023} and machine learning
approaches (e.g. \citet{dong2024}, \citet{lin2025}), which also
underscores the need for better understanding of sea-ice dynamics, and
drivers of its variability.

Good initial conditions are generally required for a good forecast,
however, it is not entirely known to what extent accurate sea ice
initial conditions affect the quality of the forecast and at what
timescales. Exploring seasonal predictions of Arctic sea ice,
\citet{guemas2016} found that sea ice initial conditions are important
in autumn to predict summer sea ice, but the impact wasn't as dramatic
when predicting winter sea ice. \citet{day2014} also found
seasonally-varying differences in the effect of initialisation, noting
that accurate Arctic sea-ice thickness lead to improved sea ice
forecasts initialised in July but not when initialised in January.

For the Antarctic, \citet{holland2013} studied the initial-value
predictability of Antarctic sea ice in a perfect model study using the
CCSM3 model. They found that sea-ice and ocean initial conditions
provide predictive information to forecast sea-ice edge location several
months in advance and that some predictability is retained for up to two
years thanks to ocean heat content anomalies that are advected eastward.
This is in contrast with \citet{marchi2020}, who ran perfect model
experiments to argue that uncertainty in the predicted atmospheric state
and evolution is the main driver of uncertainty in Antarctic sea-ice
extent prediction on seasonal timescales, with sea-ice and ocean initial
conditions having lesser importance. More recently, \citet{morioka2022}
studied decadal forecasts of Antarctic sea ice and found that
initialising ocean and sea ice improved the correlation between
simulated and observed sea-ice concentration evolution in the
Amundsen--Bellingshausen Sea. It is hard to compare these studies since
they are based on forecasts initialised at different times of the year
and different frameworks: \citet{holland2013} ran 20 ensemble members
initialised on the 1st of January of a particular year,
\citet{marchi2020} ran forecasts from the 1st of March and 1st of
September and \citet{morioka2022} run forecasts only from the 1st of
March. \citet{marchi2020} also used a coupled ocean--sea-ice model
instead of a fully coupled model like \citet{holland2013} did.
\citet{morioka2022} used observed sea ice initial conditions and
compared with observations, while \citet{marchi2020} and
\citet{holland2013} are perfect model studies.

In October 2021 the Australian Bureau of Meteorology (BoM) upgraded the
Australian Community Climate and Earth System Simulator -- Seasonal
(ACCESS-S) from version S1 to S2. While the base model remained the
same, the change in version was focused on using ocean, sea-ice and land
initial conditions generated by the BoM instead of depending on the UK
Met Office. Crucially, compared to ACCESS-S1, ACCESS-S2 does not
assimilate sea-ice observations, so sea ice is only affected by the
ocean and atmospheric data assimilation via the coupled integration.

Since model configuration is identical between ACCESS-S1 and ACCESS-S2,
they form a sort of ``natural experiment'' where the same forecasting
model was run over a long period of time with multiple ensemble
forecasts initialised throughout the year, with the only difference
being the initial conditions. This provides an opportunity to test the
effect of sea ice initial conditions on the forecast of sea-ice
concentrations and the climate.

In this study we compare sea ice hindcasts produced by ACCESS-S1 and
ACCESS-S2. We focus on seasonality of errors and biases and the effect
of the data assimilation system. This comparison will also inform future
work with the prediction system as a research tool to better understand
the dynamics and variability of the Antarctic sea ice and its impacts on
the climate system as well as to explore the potential of using its sea
ice forecasts for decision-making. The work will also serve as a
benchmark for future prediction systems to attempt to improve upon.

\section{Data and methods}\label{data-and-methods}

\subsection{ACCESS-S2}\label{s2}

ACCESS-S2 \citep{wedd2022} is the Bureau of Meteorology's seasonal
forecast system which became operational in October 2021, replacing the
ACCESS-S1 system \citep{hudson2017}. The model components of both
ACCESS-S2 and ACCESS-S1 are identical with the same numbers of levels
and resolution. They consist of the Global Atmosphere 6.0 (GA6)
\citep{williams2015, waters2017}, the Unified Model's Global Land 6.0
\citep{best2011, waters2017}, NEMO Global Ocean 5.0
\citep{gurvan2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
\citet{rae2015}{]}. The atmosphere has a N216 horizontal resolution
(\textasciitilde60 km in the mid-latitudes) with 85 vertical levels. The
land model uses the same horizontal grid as the atmosphere with four
soil levels. The ocean component has a nominal horizontal resolution of
1/4º with 75 vertical levels. The sea-ice component, based on CICE
version 4.1, has the same resolution as the ocean component and five
sea-ice thickness categories as well as an open water category.

Both systems take atmospheric initial conditions derived from
ERA-interim \citep{dee2011} for their hindcasts. The main difference
between the hindcasts of the two systems are the ocean and sea ice
initial conditions. ACCESS-S1's ocean and sea ice initial conditions
come from the Met Office FOAM system, which uses a multivariate,
incremental three-dimensional variational (3D-Var),
first-guess-at-appropriate-time (FGAT) data assimilation scheme
\citep{waters2015} and assimilates sea surface temperature (SST), sea
surface height (SSH), in situ temperature and salinity profiles, and
satellite observations of sea-ice concentration using the EUMETSAT
OSISAF product described in the next section. ACCESS-S2, on the other
hand, is initialised from ocean conditions generated by the BoM weakly
coupled ensemble data assimilation scheme described in \citet{wedd2022}.
This scheme uses an optimal interpolation method and assimilates
temperature and salinity profiles from EN4 \citep{good2013}. SSTs are
nudged to Reynolds OISSTv2.1 \citep{reynolds2007} in areas where SSTs
are over 0ºC and Sea Surface Salinity is weakly nudged to the World
Ocean Atlas 2013 climatology \citep{zweng2013}.

Of most relevance for this work, sea-ice concentrations are not
assimilated in ACCESS-S2. Assimilation cycles are performed daily. The
coupled model runs for 24 hours initialised from the previous cycle.
Then the restart file fields of the ocean component are used as first
guess in the data assimilation cycle and the innovations are used to
build the next ocean initial conditions for the following cycle. The
atmosphere fields from that daily integration are not used and instead
the model atmosphere is initialised using ERA-Interim. The sea ice
initial conditions for the next cycle are the unaltered output of the
previous daily integration. Then the cycle starts again and the coupled
model runs for another 24 hours. During this integration the sea-ice
component is affected by the ocean innovations and the new atmosphere
initial conditions via the coupler.

The ACCESS-S1 hindcast set is made up of nine members created by
perturbing the atmospheric fields only with a random field perturbation
\citep{hudson2017} and runs for 217 days for the period 1990--2012
initialised at the first of every month. The ACCESS-S2 hindcast set used
in this study runs for the period 1981--2018. Each hindcast consists of
nine ensemble members built from three consecutive 3-member forecasts
created in the same manner as ACCESS-S1 members initialised at the first
of every month and the two previous days and run for 279 days. We
analyse the ensemble mean hindcasts unless otherwise specified.

Anomalies for each hindcast set are taken with respect to their own
climatology specific to each initialisation date and forecast lead time,
for the period 1990--2012. This serves as a first-order correction of
model bias and drift. For monthly means, we define ``0 lead time
months'' as the monthly mean forecast of the same month of
initialisation.

Besides sea-ice concentration, we also analyse mean sea-ice thickness,
which we compute as total sea-ice volume divided by total sea-ice area.

\subsection{Verification datasets}\label{verification-datasets}

For verification we use satellite-derived sea-ice concentration, which
estimates the proportion of each grid area that is covered with ice.
Datasets derived using different algorithms and satellite platforms,
each have their own biases and uncertainties. Estimates of inter-product
uncertainty of sea-ice extent (SIE, defined here as the total region of
the Southern Ocean with at least 15\% sea-ice cover) is of the order of
0.5 million \(km^2\) \citep{meier2019}. As will be shown below, this
spread is minimal compared with the typical errors in the ACCESS-S2 and
ACCESS-S1 forecasts, so the overall conclusions of this study are
independent of the verification dataset used.

We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
as the primary sea-ice verification dataset. It takes the maximum value
of the NASA Team \citep{cavalieri1984} and NASA Bootstrap
\citep{comiso2023} sea-ice concentration products to reduce their low
concentration bias \citep{meier2014, meier2021}. Both source algorithms
use data from the Scanning Multichannel Microwave Radiometer (SMMR) on
the Nimbus-7 satellite and from the Special Sensor Microwave/Imager
(SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP)
-F8, -F11, and -F13 satellites. The data has a spatial resolution of 25
by 25 km and daily from 1978 onwards.

The European Organisation for the Exploitation of Meteorological
Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
\citep{OSISAF} based on the SSMIS sensor is another satellite-derived
sea-ice concentration product. It is based on mostly the same sensors as
the NOAA CDR but computed independently using different algorithms.
Figures prepared with this dataset are provided in the supplementary
material and do not differ significantly from the ones prepared using
CDR.

\subsection{Error measures}\label{error-measures}

For evaluation purposes, we use a series of measures.

Sea-ice extent is defined as the area of the ocean at least 15\% covered
by sea-ice. This threshold is motivated by the limitations in satellite
retrieval, which is increasingly unreliable for lower sea-ice
concentrations \citep{cavalieri1991}.

Pan-Antarctic (net) sea-ice extent serves as a hemispheric measure of
the amount of sea ice, but it does not take into account the spatial
distribution. A model could have a relatively accurate extent of the net
ice but with different regional distributions. To account for location
errors, we computed the Root Mean Squared Error (RMSE) of grid-point
sea-ice concentration anomalies.

We compute RMSE as the square root of the area-averaged squared
differences between grid-point forecasted and observed sea-ice
concentration anomalies. We compute a pan-Antarctic RMSE by averaging
over the whole NOAA/NSIDC CDRV4 Southern Hemisphere domain, and also a
zonally-varying RMSE computed over 15 longitude slices 24° wide around
Antarctica.

All error measures were computed on the NOAA/NSIDC CDRV4 domain grid, to
which model output was bilinearly interpolated. Note that the ACCESS
CICE model grid has resolution between two and three times higher than
NOAA/NSIDC CDRV4.

Forecasts errors are also compared with hypothetical forecasts based on
the persistence of anomalies and on climatology. The persistence
forecast is generated by extending the observed anomalies in time.

As a measure of forecast improvement over the hypothetical forecast, we
use the skill score \citep{murphy1985}, defined as

\[
S = 1 - \frac{RMSE_{f}}{RMSE_{r}}
\]

Where \(RMSE_{f}\) is the RMSE of the forecast, \(RMSE_{r}\) is the RMSE
of the reference forecast. Negative skill score indicates that the
forecast is worse than the reference forecast while positive values
indicate an improvement. A perfect forecast, would have zero RMSE and
thus a skill score of 1.

\section{Results and discussion}\label{results-and-discussion}

\subsection{Bias}\label{bias}

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}}

}

\caption{\label{fig-hindcast-extent}Row a: Pan-Antarctic mean sea-ice
extent for all hindcasts initialised on the first of each calendar month
for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2; purple).
Observed mean sea-ice extent in each corresponding hindcast period is
shown in black. Row b: Mean differences between the forecast and the
observed values. Circles represent the initial conditions at the start
of forecasts (i.e., the first of every month), and triangles represent
the mean values at the end of forecasts (i.e., the longest possible lead
time).}

\end{figure}%

Figure~\ref{fig-hindcast-extent} shows mean sea-ice extent of the
ACCESS-S1 and ACCESS-S2 hindcasts (row a) and their differences from
mean sea-ice extent of NOAA/NSIDC CDRV4 (row b). Mean extent at the
first of every month is indicated with circles for the initial
conditions and with triangles for the longest lead time possible for
each model (between 274 and 277 days for ACCESS-S2 and between 213 and
216 days for ACCESS-S1). At this long lead time, information of the
initial conditions is essentially lost and the forecast reverts to each
model's preferred equilibrium state.

ACCESS-S2 initial conditions (circles in Fig.~\ref{fig-hindcast-extent}
column 2) show an overall negative bias, especially in the late
summer-early autumn, while ACCESS-S1 initial conditions (circles in
Fig.~\ref{fig-hindcast-extent} column 1) are very close to observations,
as expected from the assimilation of sea-ice observations to produce the
initial conditions of ACCESS-S1. Both systems' equilibrium states
(triangles) show negative biases of sea-ice extent, particularly in the
growth phase of late-autumn and winter months. This is due primarily to
the melt season being longer than in observations and with faster melt
between January and March and the growing seasons being shorter with
slower growth during March and April. This is then followed by faster
growth between May and July (Figure~\ref{fig-mean-growth}). Many sea-ice
models exhibit this systematic underestimation during the sea-ice
minimum and early freezing season \citep{massonnet2023}, which could
indicate problems in the representation of thermodynamics in the model
\citep{zampieri2019}. It is also not surprising that both forecasting
systems converge to a similar equilibrium state because they share the
same model formulation.

The difference between the initial conditions (circles) and the model
equilibrium state (triangles) can be mostly attributed to the effect of
data assimilation, which in ACCESS-S2 is due solely to the coupling of
sea-ice with the atmosphere and the ocean. From May to October, in
ACCESS-S2 circles are closer to observations than the triangles are,
indicating that the information from the ocean and atmosphere data
assimilation is affecting sea ice and improving the initial conditions.
During these months, ACCESS-S1 can overestimate the sea-ice extent at
short lead time. For the rest of the year circles are overlaid with
triangles in ACCESS-S2, indicating that the ocean and atmosphere data
assimilation is not affecting sea ice and that this component of the
model is virtually free-running.

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}}

}

\caption{\label{fig-mean-growth}Mean daily sea-ice extent growth
(\(10^6 km^2/day\)) in ACCESS-S1 (green) and ACCESS-S2 (purple)
hindcasts and observations (black), computed as the mean daily
differences in sea-ice extent between each date and the next for each
forecast month. Values are smoothed with a 11-day running mean.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1-1.pdf}}

}

\caption{\label{fig-bias-1}Ensemble mean difference between monthly
sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month
lead time (monthly mean values forecasted from the forecast initialised
at the first of the month) and observations.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-2-1.pdf}}

}

\caption{\label{fig-bias-2}Same as Figure~\ref{fig-bias-1} but for
ACCESS-S1}

\end{figure}%

To further understand the bias in ACCESS-S2, Figure~\ref{fig-bias-1}
shows spatial patterns of the differences of monthly mean sea-ice
concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
shortest monthly lead time. From October to May, the model
underestimates sea-ice concentrations in most regions except for the
inner Weddell Sea in April and May, where sea-ice concentrations
saturate to 1 both in the observations and forecasts. In winter, the
differences are limited to a narrow band around the sea-ice edge with
slight positive biases in the African sector of East Antarctica and
negative biases around the Indian Ocean sector which partially
compensate, resulting in the near-zero extent bias seen in those months
(Figure~\ref{fig-hindcast-extent}).

ACCESS-S1 has a comparatively smaller overall bias
(Figure~\ref{fig-bias-2}). The largest values are found between April
and June, when the faster growth results in large positive bias along
the sea-ice edge, and in January, when the faster melt leads to large
negative bias in the Weddell and Admunsen seas.

\FloatBarrier

\subsection{RMSE}\label{rmse}

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-1.pdf}}

}

\caption{\label{fig-extent-anom}Monthly mean sea-ice extent anomalies of
the observations (black) and forecasts from ACCESS-S1 (right column;
purple) and ACCESS-S2 (left column; green) at lead times of 0, 2, 4, and
6 months. The RMSE and correlation during the overlapping period of
ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) is shown on the top left
and bottom left of each panel respectively.}

\end{figure}%

Figure~\ref{fig-extent-anom} shows monthly sea-ice extent anomalies
forecasted at selected lead times. Compared with ACCESS-S1, ACCESS-S2
anomaly forecasts are relatively poor (large RMSE) even for the first
month (lead time 0), whereas ACCESS-S1 forecasts stay relatively skilful
even at a lead time of three months. ACCESS-S2 shows much larger
interannual variability than observations, with dramatic lows between
1995 and 2007, and highs between 2007 and 2015.

Unexpectedly, for ACCESS-S2, RMSE improves with lead time, even though
the correlation degrades with lead time. This is puzzling behaviour that
goes contrary to what is usually seen in prediction models. The
explanation seems to be the mentioned increased interannual variability.
Figure~\ref{fig-extent-sd} shows the interannual standard deviation of
monthly sea-ice extent of the forecasts as a function of lead time
compared with observations. ACCESS-S1 standard deviation lies within the
observed standard deviation regardless of lead time, while ACCESS-S2
standard deviation is more than twice that of observations at zero lead
time and only approaches the observed value at 9 month lead time for
most months.

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-sd-1.pdf}}

}

\caption{\label{fig-extent-sd}Interannual standard deviation with 95\%
confidence interval of monthly mean sea-ice extent forecasted for each
month divided by that month's sea-ice extent observation standard
deviation. ACCESS-S1 and ACCESS-S2 at different lead times. Each panel
indicates the target month. Note the reverse horizontal axis.}

\end{figure}%

ACCESS-S2 forecasts of sea-ice extent anomalies seem to align moderately
well with observations (leading to moderately high correlation) but
their magnitude is overestimated (leading to large errors). This could
be caused by ACCESS-S2 sea ice being much more sensitive to atmospheric
and oceanic forcing, perhaps due to lower thickness.

As an example, Figure~\ref{fig-forecast-example} shows sea-ice
concentration anomalies (top row) and sea-ice thickness and the
difference between the two models (bottom row) for 2 May 2008
initialised one day prior; being that close to initialisation date,
these are very approximately the initial conditions. ACCESS-S1 sea-ice
concentrations anomalies are very close to observations as expected from
the system assimilating these data. ACCESS-S2 sea-ice concentration
anomalies, which are not assimilated, are not as close, but the
large-scale pattern is aligned with observations. The system simulates
large positive anomalies in the Weddell and Ross seas and slight
negative anomalies in the Amundsen and Bellingshausen seas. The fact
that ACCESS-S2 can simulate this pattern without assimilating sea-ice
data suggests that atmospheric and oceanic forcing was the dominant
factor. However, the magnitude of the sea-ice anomalies is too big. It
is plausible that this is due to the thinner ice simulated by ACCESS-S2
(bottom row).

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-forecast-example-1.pdf}}

}

\caption{\label{fig-forecast-example}ACCESS-S1 and ACCESS-S2 hindcasts
for 2 May 2008 at one day lead time. Top row shows sea-ice concentration
anomalies forecasted by each system and the observations. Bottom row
shows forecasted sea-ice thickness and the difference between ACCESS-S1
and ACCESS-S2.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-thickness-1.pdf}}

}

\caption{\label{fig-mean-thickness}Mean and 95\% interval of monthly
mean sea-ice thickness for ACCESS-S1 and ACCESS-S2 at different lead
times. Each panel indicates the target month. Note the reverse
horizontal axis.}

\end{figure}%

Extending beyond the one case in Figure~\ref{fig-forecast-example},
Figure~\ref{fig-mean-thickness} shows monthly mean sea-ice thickness as
a function of lead time for ACCESS-S1 and ACCESS-S2. Supporting the idea
that thinner ice is what causes the increased extent variability in
ACCESS-S2, this system simulates thinner sea-ice compared to ACCESS-S1
overall at almost all lead times and in all months except for summer at
short lead times (Dec-Jan, 0-1 months; Feb-Mar, 0-2 months). However, in
both systems, forecasted sea-ice is thicker at shorter lead times and
then decreases, particularly in the summer months. If thinner ice were a
sufficient cause of increased variability, then we would expect
variability to increase with lead time in both forecasting systems.

The fact that ACCESS-S1 and ACCESS-S2 share the same model configuration
and that the increased variability is more extreme at short lead times
(Fig.~\ref{fig-extent-sd}) suggests that the data assimilation procedure
is partly to blame. It is possible that sea-ice in the ACCESS-S2 system
is left in an unbalanced state after assimilating atmospheric and
oceanic data but not sea-ice data, leading to large responses that are
amplified by the thin ice in the initial states which then subside at
longer lead times when the model is balanced.

To assess ACCESS-S2 forecasts in more detail, we compute error measures
for all hindcasts started on the 1st of every month.

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse-1.pdf}}

}

\caption{\label{fig-rmse}Mean RMSE of sea-ice concentration anomalies as
a function of forecast lead time for all forecasts initialised on the
first of each month compared with a reference forecast of persistence of
anomalies (black) and climatology (gray). Only the first 120 days are
shown. In parenthersis, the shortest time at which ACCESS-S1 and
ACCESS-S2 mean RMSE is not statistically different at the 99\%
confidence level.}

\end{figure}%

Figure~\ref{fig-rmse} shows the mean RMSE of sea-ice concentration
anomalies for ACCESS-S1 and ACCESS-S2 hindcasts compared against
persistence and climatological forecasts used as a benchmark. Due to
errors in the initial conditions, it is expected that persistence
forecasts would be better than the model forecasts at very short lead
times, but that the persistence forecast errors would grow faster and
may eventually surpass the model forecast errors. The black line shows
that the persistence forecast error indeed grows rapidly and reaches its
maximum in about 30 days for most months except for February, when it
grows much slower. The ACCESS-S1 forecast errors grow slower than
persistence forecast errors and remain lower after less than 10 days on
average. The ACCESS-S2 forecast error starts high in all months and is
lower than the persistence forecast error after more than 15 days in
most months except for forecast initialised in February, when it takes
80 days.

At longer lead times, it is more appropriate to compare errors with the
climatological forecast error. The lead time at which ACCESS-S1 forecast
error is higher than the climatological forecast error varies between
more than 60 and less than 20 days depending on forecast initialisation
month with the minimum in June. ACCESS-S2 forecasts never have lower
error than climatology, on the other hand, except marginally in October
forecasts.

Figure~\ref{fig-lead-time-window} summarises the lead time window in
which each hindcast is better than both the persistence forecast and the
climatological forecast as a function of forecast month. ACCESS-S1
forecasts have a wider lead time window in the summer than the other
seasons and is not better than both benchmarks at forecasting June
sea-ice concentration anomalies. Forecasts initialised in May and June
are particularly poor, and July cannot be forecasted better than the
benchmarks. This is consistent with the mid-winter loss of
predictability observed by \citet{libera2022}, who attributed it to deep
warm water entraining into the mixed layer.

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-lead-time-window-1.pdf}}

}

\caption{\label{fig-lead-time-window}Minimum lead time at which each
forecast's mean RMSE becomes larger than the lower bound of the 95\%
confidence interval of persistence forecast RMSE (black lines) and
maximum lead time at which each forecast's mean RMSE remains lower than
the lower bound of the 95\% confidence interval of climatological
forecast RMSE (gray lines). Green shading indicates the window where
forecasts outperform both persistence (lead times longer than black
line) and climatology (lead times shorter than gray line). Text labels
show the date corresponding to the maximum lead time at which each
forecast outperforms climatology.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-1.pdf}}

}

\caption{\label{fig-rmse_lon-1}RMSE skill score of ACCESS-S1 forecasts
with climatological forecast as reference computed on 15 meridional
slices 24º wide as a function of lead time and longitude. Antarctica's
coastline is shown at the bottom of each panel for reference.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-1.pdf}}

}

\caption{\label{fig-rmse_lon-2}Same as Figure~\ref{fig-rmse_lon-1} but
for ACCESS-S2}

\end{figure}%

To analyse the spatial distribution of the model error, we computed the
RMSE of zonal mean sea-ice concentration anomalies on 15 slices of 24°
longitude span for each forecasting system. We control for some areas
being naturally easier to forecast than others by computing the RMSE
skill score with the climatological forecast RMSE as reference.

For ACCESS-S1 forecasts (Figure~\ref{fig-rmse_lon-1}), skill tends to be
lower off the coast of Eastern Antarctica even at short lead times; for
instance, the skill score for forecasts initialised in May and June are
negative between 0° and 120°E even at almost zero lead time. This
mirrors \citet{libera2022} findings of a ``winter predictability
barrier'', although they focus on the Weddell sea and here we show that
the effect seems to be stronger more to the East. In western Antarctica
there is a hint of easterly-propagating skill in forecasts initialised
in February and March. This is consistent with \citet{holland2013}
findings that memory of sea-ice anomalies are stored in ocean heat
content anomalies that are transported east by the Antarctic Circumpolar
Current.

ACCESS-S2 forecasts (Figure~\ref{fig-rmse_lon-2}) also have lower skill
over Eastern Antarctica. From July to December even though the
pan-Antarctic average skill is negative at all lead times
(Fig.~\ref{fig-lead-time-window}), it is positive for up to a month in
Western Antarctica. Since oceanic and atmospheric forcing is the only
source of information, this suggests that sea-ice in this region is
particularly sensitive to oceanic and atmospheric forcing and suggests a
role of the Pacific-South American mode and the Amundsen Sea Low to
shape sea-ice concentration anomalies. The fact that this is evident in
the months in which ENSO teleconnections are more important for
atmospheric circulation also suggest the influence of tropical Pacific
variability. February and March are the only two months that can be
forecasted with marginally positive skill in large regions.

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-initial-spread-1.pdf}}

}

\caption{\label{fig-initial-spread}Decomposition of forecast error
spread at 1, 5 and 30 days lead time for ACCESS-S1 and ACCESS-S2
hindcasts across initialization months. The left column shows the mean
standard deviation of RMSE errors across ensemble members, while the
right column shows the standard deviation of the ensemble mean RMSE
error and the spread of the persistence and climatology forecasts
errors.}

\end{figure}%

In Figure~\ref{fig-rmse} the mean error was shown.
Figure~\ref{fig-initial-spread} column 1 shows the mean standard
deviation of errors among ensemble members at various lead times. At one
day lead time (Fig.~\ref{fig-initial-spread} a.1) ACCESS-S2 has a
slightly larger spread than ACCESS-S1 due to the way that ensemble
members are generated. ACCESS-S1 ensemble members are generated by
adding random field perturbations to the atmosphere only, which then are
transferred to the other components via the coupled simulation
\citep{hudson2017}. With this scheme, ensemble members are all but
guaranteed to be underdispersed in the ocean and sea-ice components. The
time-lag ensemble used for ACCESS-S2 ensures greater spread. This
difference is gone after about just two days, and both systems have a
comparable spread in ensemble member error afterwards
(Fig.~\ref{fig-initial-spread} b1 and c1).

Figure~\ref{fig-initial-spread} column 2, on the other hand, shows the
standard deviation of ensemble mean error of each hindcast and the
persistence forecast. At one day lead time, ACCESS-S2 ensemble mean
error standard deviation is much larger than ACCESS-S1's, which in turn
is comparable to the persistence forecast error standard deviation. At
longer lead times, the spread of ACCESS-S1 and persistence forecast
standard deviation increases to eventually be comparable to ACCESS-S2
and the standard deviation in climatological forecast errors. ACCESS-S2
error standard deviation is fairly independent of lead time and similar
to the climatological forecast error standard deviation at all lead
times.

\FloatBarrier

\subsection{Conclusions}\label{conclusions}

Sea ice forecasts from the ACCESS-S2 system show a significant low
extent bias, particularly during late summer and early autumn. This bias
is attributed to a faster and longer melt season between January and
March, and slower growth between March and April. This underestimation
during the minimum and early freezing season is a common issue in many
seasonal-to-subseasonal (S2S) systems, suggesting potential problems
either with the model's thermodynamic representation or with short wave
radiation forcing, as shown in other climate models
\citep{zampieri2019, roach2020}. Even though ACCESS-S2 shares the same
model formulation with ACCESS-S1, the latter does not suffer from this
bias, indicating that assimilating sea-ice concentrations successfully
corrects for the negative bias that exists in the free-running model.

Ensemble spread grows quickly even when perturbations are only
implemented in the atmosphere component (in ACCESS-S1), indicating that
sea-ice is indeed responding quickly to atmospheric perturbations.
However, our analysis suggests that the atmosphere and ocean data
assimilation implemented in ACCESS-S2 is only effectively influencing
sea ice initial conditions from June to October, while the rest of the
year, the sea-ice component runs virtually free, reverting to its biased
equilibrium state.

Analysis of the error spread shows that ACCESS-S2 initial conditions
from December to May not only have large errors, but that the initial
error spread is very large compared with ACCESS-S1. This spread is not
due to the perturbation scheme, since the mean error variance for
individual forecasts is low and comparable with ACCESS-S1. Instead, it
is due to large variance of the mean error of individual forecasts,
which is comparable to the climatology spread. This is further evidence
that individual initial conditions are not being affected by the data
assimilation scheme.

Although ACCESS-S1 only assimilates sea-ice concentration, it is clear
that sea-ice thickness is also affected. ACCESS-S1 simulates
significantly thicker ice than ACCESS-S2 and in both systems sea-ice is
thicker at shorter lead times than at longer lead times. Both the
explicit data assimilation in ACCESS-S2 and the effects of atmospheric
and oceanic data assimilation in ACCESS-S1 might be nudging simulated
sea ice to be thicker than the model equilimbrium state.\\
We believe that the thinner sea ice in ACCESS-S2 is a partial cause for
the large sea-ice extent variance, but other mechanisms, such as
unbalanced initial conditions might be important.

Given that ACCESS-S2 sea-ice extent is essentially not initialised by
the observations, comparing its forecasts with those of ACCESS-S1 allows
us to estimate the time-scale for which initial conditions are
important. We find that initial conditions affect Antarctic sea ice
forecasts in the order of a few months, but that effect is seasonally
dependent. January to April initial conditions improve forecasts for up
to 3 months with, February initial conditions in particular are shown to
be crucial for determining sea-ice evolution at least up to May. Arctic
sea ice forecasts also show greater sensitivity to initial conditions in
boreal summer, compared with boreal winter \citep{day2014, bunzel2016},
suggesting a similar mechanism might be playing a role.

Forecasts initialised in winter have very little skill and ACCESS-S1 and
ACCESS-S2 forecast errors are not statistically different after just 2
weeks. This is consistent with \citet{libera2022}'s finding of a
``winter predictability barrier'' in the Weddell sea, although they
describe the barrier as sharp loss of predictability in July, and here
we find a gradual reduction in skill compared with climatology around
June. This difference might be due to our use of pan-Antarctic RMSE,
since our regional analysis indicates that the degraded skill is most
dramatic in the King Haakon sea.

These findings have important implications for both operational
forecasting, model development and predictability studies. For
operational centers, our results suggest that efforts to improve sea-ice
data assimilation should prioritize the summer and autumn months when
initial conditions have the greatest impact on forecast skill.
Additionally, the substantial bias in ACCESS-S2 highlights the need for
improved model physics, particularly in the representation of sea-ice
thermodynamics and radiation processes. Crucially, our results suggest
dramatic seasonal variations in sea-ice predictability. Future studies
should therefore use initial conditions through the whole year rather
than focusing on a limited number of initialisation dates.

\section{References}\label{references}

\renewcommand{\bibsection}{}
\bibliography{references.bib}

\clearpage

\appendix  
\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}

\section{Supplementary figures}\label{supplementary-figures}

The following are the same figures from the main paper but using the OSI
dataset instead of CDR.

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-hindcast-extent-osi-1.pdf}}

}

\caption{\label{fig-hindcast-extent-osi}Row a: Pan-Antarctic mean
sea-ice extent for all hindcasts initialised on the first of each
calendar month for ACCESS-S1 (column 1; green) and ACCESS-S1 (column 2;
purple). Observed mean sea-ice extent in each corresponding hindcast
period is shown in black. Row b: Mean differences between the forecast
and the observed values. Circles represent the initial conditions at the
start of forecasts (i.e., the first of every month), and triangles
represent the mean values at the end of forecasts (i.e., the longest
possible lead time).}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-mean-growth-osi-1.pdf}}

}

\caption{\label{fig-mean-growth-osi}Mean daily sea-ice extent growth
(\(10^6 km^2/day\)) in ACCESS-S1 (green) and ACCESS-S2 (purple)
hindcasts and observations (black), computed as the mean daily
differences in sea-ice extent between each date and the next for each
forecast month. Values are smoothed with a 11-day running mean.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-1-osi-1.pdf}}

}

\caption{\label{fig-bias-1-osi}Ensemble mean difference between monthly
sea-ice concentration of ACCESS-S2 ensemble mean forecast at 0-month
lead time (monthly mean values forecasted from the forecast initialised
at the first of the month) and observations.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-bias-2-osi-1.pdf}}

}

\caption{\label{fig-bias-2-osi}Same as Figure~\ref{fig-bias-1} but for
ACCESS-S1}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-anom-osi-1.pdf}}

}

\caption{\label{fig-extent-anom-osi}Monthly mean sea-ice extent
anomalies of the observations (black) and forecasts from ACCESS-S1
(right column; purple) and ACCESS-S2 (left column; green) at lead times
of 0, 2, 4, and 6 months. The RMSE and correlation during the
overlapping period of ACCESS-S1 and ACCESS-S2 hindcasts (1990--2013) is
shown on the top left and bottom left of each panel respectively.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-extent-sd-osi-1.pdf}}

}

\caption{\label{fig-extent-sd-osi}Interannual standard deviation with
95\% confidence interval of monthly mean sea-ice extent forecasted for
each month divided by that month's sea-ice extent observation standard
deviation. ACCESS-S1 and ACCESS-S2 at different lead times. Each panel
indicates the target month. Note the reverse horizontal axis.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse-osi-1.pdf}}

}

\caption{\label{fig-rmse-osi}Mean RMSE of sea-ice concentration
anomalies as a function of forecast lead time for all forecasts
initialised on the first of each month compared with a reference
forecast of persistence of anomalies (black) and climatology (gray).
Only the first 120 days are shown. In parenthersis, the shortest time at
which ACCESS-S1 and ACCESS-S2 mean RMSE is not statistically different
at the 99\% confidence level.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-lead-time-window-osi-1.pdf}}

}

\caption{\label{fig-lead-time-window-osi}Minimum lead time at which each
forecast's mean RMSE becomes larger than the lower bound of the 95\%
confidence interval of persistence forecast RMSE (black lines) and
maximum lead time at which each forecast's mean RMSE remains lower than
the lower bound of the 95\% confidence interval of climatological
forecast RMSE (gray lines). Green shading indicates the window where
forecasts outperform both persistence (lead times longer than black
line) and climatology (lead times shorter than gray line). Text labels
show the date corresponding to the maximum lead time at which each
forecast outperforms climatology.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-1-osi-1.pdf}}

}

\caption{\label{fig-rmse_lon-1-osi}RMSE skill score of ACCESS-S1
forecasts with climatological forecast as reference computed on 15
meridional slices 24º wide as a function of lead time and longitude.
Antarctica's coastline is shown at the bottom of each panel for
reference.}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-rmse_lon-2-osi-1.pdf}}

}

\caption{\label{fig-rmse_lon-2-osi}Same as Figure~\ref{fig-rmse_lon-1}
but for ACCESS-S2}

\end{figure}%

\begin{figure}[!htb]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{access-ice_files/figure-pdf/fig-initial-spread-osi-1.pdf}}

}

\caption{\label{fig-initial-spread-osi}Decomposition of forecast error
spread at 1, 5 and 30 days lead time for ACCESS-S1 and ACCESS-S2
hindcasts across initialization months. The left column shows the mean
standard deviation of RMSE errors across ensemble members, while the
right column shows the standard deviation of the ensemble mean RMSE
error and the spread of the persistence and climatology forecasts
errors.}

\end{figure}%

\end{document}